---
title: "adaptiveFTS"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    css: style.css
    fig_width: 8
    fig_height: 3

vignette: >
  %\VignetteIndexEntry{adaptiveFTS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
echo = TRUE,
message = FALSE,
warning = FALSE,
collapse = TRUE,
comment = "#>"
)
```

```{r setup}
library(adaptiveFTS)
library(data.table)
library(ggplot2)
```

This vignette gives an overview of the adaptive estimation of the mean and the
(auto)covariance functions, and of the adaptive prediction of a curve by the Best
Linear Unbiased Predictor (BLUP). See the following references for the
methodology.

1. **Hassan Maissoro, Valentin Patilea, and Myriam Vimond.** *Adaptive Estimation for Weakly Dependent Functional Time Series.* Journal of Time Series Analysis, 2025. [doi:10.1111/jtsa.70006](https://doi.org/10.1111/jtsa.70006).

2. **Hassan Maissoro, Valentin Patilea, and Myriam Vimond.** *Adaptive Prediction for Functional Time Series.* 2026. arXiv:2609.xxxxx.


# The data

The unit of observation is a curve. The data is a collection of $N$ curves $\{X_1, \ldots, X_N\}$ that are realizations of a process $X$. For each $1\leq n \leq N$, the trajectory (or curve) $X_n$  is observed at the domain points $\{T_{n,i}, 1\leq i \leq  M_n\}\subset I$, with additive noise. The data points associated with $X_n$ consist of  the pairs  $(Y_{n,i} , T_{n,i} ) \in\mathbb R \times I $, where
$$
Y_{n,i} = X_n(T_{n,i}) +  \sigma(T_{n,i})\varepsilon_{n,i}, \qquad 1\leq n \leq N,  \; \; 1\leq i \leq M_n.
$$
The data generating process satisfies the following assumptions.

- The series $\{X_n\}$ is a (strictly) stationary $\mathcal H-$valued series.

- The $M_1, \dotsc, M_N$ are random draws of an integer  variable $M\geq 2$, with expectation $\lambda$.

- Either all the $T_{n,i}$  are independent copies of a  variable $T\in I$ which admits a strictly positive density $g$ over $I$ (independent design case), or the $T_{n,i}$, $1\leq i \leq \lambda=M_n$, are the points of the same equidistant grid of $\lambda$ points in $I$ (common design case).

- The $\varepsilon_{n,i}$  are independent copies of a centered error variable $\varepsilon$ with unit variance, and  $\sigma^2(\cdot)$ is a Lipschitz continuous function.

- The series $\{X_n\}$ and the copies of  $M$, $T$ and $\varepsilon$ are mutually independent.

Every estimator takes the data through `format_data()`, which reshapes it into a
`data.table` with three columns: `id_curve`, `tobs` and `X`. The curve index
defines the order of the series, so the curves must be given in chronological
order. The domain is $I = [0, 1]$.

The package ships `data_far`, a sample drawn from a functional autoregressive
process of order one, FAR(1).

# Sample of FTS generation

`simulate_far()` draws a FAR(1) sample. The curves are built on an integration
grid of `n_int_grid` points, then observed at `M_distribution` random points per
curve drawn from `t_distribution` (independent design) or at the common grid
`t_common` (common design). The roughness of the sample paths is set by the Hurst
function `hurst_fun` and the squared Hölder constant `L2`, and the first `n_burnin` curves
are dropped so that the returned sample is close to stationary.

```{r}
set.seed(42)
dt_sim <- simulate_far(
  N = 50L, lambda = 40L,
  design = "random",
  M_distribution = rpois,
  t_distribution = runif,
  t_common = NULL,
  hurst_fun = hurst_logistic,
  L2 = 4,
  far_kernel = function(s, t) 9 / 4 * exp(-(t + 2 * s) ** 2),
  far_mean = function(t) 4 * sin(1.5 * pi * t),
  n_int_grid = 100L,
  n_burnin = 100L,
  remove_burnin = TRUE)

head(dt_sim)
```

The packaged `data_far` was drawn the same way. Its curves look like this.

```{r}
data("data_far")

ggplot(data = data_far[, list("id_curve" = as.factor(id_curve), "t" = tobs, X)],
       mapping = aes(x = t, y = X, group = id_curve)) +
  geom_line(colour = "#1B4F72", alpha = 0.3) +
  labs(x = "t", y = "X(t)") +
  theme_minimal()
```

The estimation sections below run on the first 60 curves, to keep this vignette
quick to build.

```{r}
dt_sub <- data_far[id_curve <= 60]
```

# Estimation of local regularity parameters

`estimate_locreg()` estimates the local Hölder exponent $H_t$ and the squared
Hölder constant $L_t^2$ at each point of `t`. Together they drive the bandwidth
chosen by every estimator below: the rougher the process at $t$, the smaller the
bandwidth.

```{r}
t0 <- seq(0.1, 0.9, len = 9)

dt_locreg <- estimate_locreg(data = dt_sub,
                             idcol = "id_curve",
                             tcol = "tobs",
                             ycol = "X",
                             t = t0,
                             Delta = NULL,
                             presmooth_bw = NULL,
                             kernel_name = "epanechnikov",
                             center = TRUE)

dt_locreg
```

Every estimator output carries `summary()` and `plot()` methods.

```{r}
summary(dt_locreg)
plot(dt_locreg)
```

# Estimation of the mean function

The adaptive mean estimator selects its bandwidth point by point, by minimising
the estimated risk over a grid of candidates. `estimate_mean_risk()` exposes that
risk, split into its bias, variance and dependence terms.

```{r}
bw_grid <- seq(0.02, 0.15, length.out = 10)

dt_mean_risk <- estimate_mean_risk(
  data = dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X",
  t = c(1/4, 1/2, 3/4), bw_grid = bw_grid, kernel_name = "epanechnikov")

plot(dt_mean_risk)
```

`estimate_mean()` runs the same selection and returns the estimate, reporting the
bandwidth used at each point in the `optbw` column.

```{r}
dt_mean <- estimate_mean(
  data = dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X",
  t = seq(0.1, 0.9, length.out = 25), bw_grid = bw_grid,
  kernel_name = "epanechnikov")

summary(dt_mean)
plot(dt_mean)
```

# Estimation of the autocovariance function

The lag-$\ell$ autocovariance is estimated at pairs of points $(s, t)$. Two
estimators are available, selected by `common_bw`.

## Using one bandwidth

With `common_bw = TRUE` a single bandwidth is selected for both arguments. This
is the estimator of Maissoro, Patilea and Vimond (2025).

```{r}
sgrid <- c(1/5, 2/5, 3/5, 4/5)
tgrid <- c(1/4, 1/2, 3/5, 3/4)

dt_autocov_1bw <- estimate_autocov(
  data = dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X",
  s = sgrid, t = tgrid, lag = 1, bw_grid = bw_grid,
  common_bw = TRUE, center_curves = TRUE, correct_diagonal = FALSE,
  kernel_name = "epanechnikov")

dt_autocov_1bw[, list(s, t, optbw_s, optbw_t, autocov)]
```

## Using two bandwidths

With `common_bw = FALSE` the risk is minimised over pairs $(h_s, h_t)$, so the
estimator adapts to the regularity at $s$ and at $t$ separately. This is the
estimator of Maissoro, Patilea and Vimond (2026). It explores the square of the
grid and therefore costs noticeably more.

```{r}
dt_autocov_2bw <- estimate_autocov(
  data = dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X",
  s = sgrid, t = tgrid, lag = 1, bw_grid = bw_grid,
  common_bw = FALSE, center_curves = TRUE, correct_diagonal = FALSE,
  kernel_name = "epanechnikov")

dt_autocov_2bw[, list(s, t, optbw_s, optbw_t, autocov)]
```

The two bandwidths differ wherever the estimated regularity differs between $s$
and $t$.

## Functional autocorrelation

`estimate_facf()` normalises the lag-$\ell$ autocovariance surfaces into a
functional analogue of the autocorrelation function of a scalar series, which is
a convenient way to read the serial dependence of the sample.

```{r}
dt_facf <- estimate_facf(
  data = dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X",
  lag.max = 3, n_grid = 8, bw_grid = bw_grid, common_bw = TRUE,
  kernel_name = "epanechnikov")

summary(dt_facf)
plot(dt_facf)
```

# Adaptive prediction of a curve

`blup_fit()` estimates every component of the adaptive BLUP that does not depend
on the prediction points, and `predict()` evaluates the predictor on a grid. The
Tikhonov regularisation parameter is selected by cross-validation when left to
`NULL`.

```{r}
data_train <- format_data(dt_sub, idcol = "id_curve", tcol = "tobs", ycol = "X")

fit <- blup_fit(data = data_train, n_cv_curves = 10L)
summary(fit)
```

`predict()` reconstructs the curve following the conditioning one. With
`horizon > 1`, each intermediate curve is fed back as the new conditioning curve
and every intermediate prediction is returned.

```{r}
dt_pred <- predict(fit, t = seq(0.05, 0.95, length.out = 40), horizon = 2L)

ggplot(dt_pred, aes(x = t, y = prediction, colour = factor(horizon))) +
  geom_line(linewidth = 0.7) +
  labs(x = "t", y = "prediction", colour = "horizon") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
