---
title: "adaptiveFTS"
output: 
  rmarkdown::html_vignette:
    toc : true
    toc_depth : 4
    number_sections: true
    theme: readable
    highlight: tango
    toc_float:
      collapsed: false
      smooth_scroll: false
    css: style.css
    fig_width: 8
    fig_height: 3

vignette: >
  %\VignetteIndexEntry{adaptiveFTS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(adaptiveFTS)
```

```{r}
library(data.table)
library(ggplot2)
library(ggpubr)
library(plotly)
library(magrittr)
library(dygraphs)
library(manipulateWidget)
library(crosstalk)
```

The vignette proposes an overview on how to perform the adaptive estimation procedure of mean and the covariance functions estimation as well as the adaptive prediction of the unobserved part of a curve using the Best Linear Unbiased Estimator (BLUP). One can refer to the following references for more ddetails.

1. **Hassan Maissoro, Valentin Patilea, and Myriam Vimond.**  *Adaptive Estimation for Weakly Dependent Functional Time Series.* 2024. Available at [arXiv:2403.13706](https://arxiv.org/abs/2403.13706).

2. **Hassan Maissoro, Valentin Patilea, and Myriam Vimond.**  *Adaptive prediction for Functional Times Series.*  2024. [Work in progress](https://hassan.maissoro.com/assets/pdf/2024-adaptive-estimation-for-functional-time-series.pdf).


# The data

The unit of observation is a curve. The data is a collection of $N$ curves $\{X_1, \ldots, X_N\}$ that are realizations of a process $X$. For each $1\leq n \leq N$, the trajectory (or curve) $X_n$  is observed at the domain points $\{T_{n,i}, 1\leq i \leq  M_n\}\subset I$, with additive noise. The data points associated with $X_n$ consist of  the pairs  $(Y_{n,i} , T_{n,i} ) \in\mathbb R \times I $, where 
$$
Y_{n,i} = X_n(T_{n,i}) +  \sigma(T_{n,i})\varepsilon_{n,i}, \qquad 1\leq n \leq N,  \; \; 1\leq i \leq M_n.
$$
The data generating process satisfies the following assumptions.

- The series$\{X_n\}$ is a (strictly) stationary $\mathcal H-$valued series.
	
- The $M_1, \dotsc, M_N$ are random draws of an integer  variable $M\geq 2$, with expectation $\lambda$.  
	
- Either all the $T_{n,i}$  are independent copies of a  variable $T\in I$ which admits a strictly positive density $g$ over $I$ (independent design case), or the $T_{n,i}$, $1\leq i \leq \lambda=M_n$, are the points of the same equidistant grid of $\lambda$ points in $I$ (common design case).
	
- The $\varepsilon_{n,i} $  are independent copies of a centered error variable $\e$ with   unit variance, and  $\sigma^2(\cdot)$ is a Lipschitz continuous function. 
	
- The series $\{X_n\}$ and the copies of  $M$, $T$ and $\varepsilon$ are mutually independent.

The package includes an example of dataset `data_far` drawn as a sample of an Functional Autoreressive Process of order one (FAR(1)). Thus, in the following, we first describe how to generate data from a FRA(1). Second, we explain the data

# Sample of FTS generation

**TODO:** Describe the procedure

```{r}
# Load the data
data("data_far")

# Plot the curve
ggplot(data = data_far[, .("id_curve" = as.factor(id_curve), "t" =  tobs, X)],
      mapping = aes(x = t, y = X, group = id_curve)) +
  geom_line(color = "#1B4F72") +
  theme_minimal()


```



# Estimation of local regularity parameters

# Estimation of the mean function

# Estimation of the autocovariance function

## Using on bandwidth

## Using two bandwidth

# Adaptive prediction of a curve




