---
title: "Prediction of a curve with FDA"
author: "Hassan Maissoro"
output: 
  html_document:
    toc : true
    toc_depth : 4
    number_sections: true
    theme: readable
    highlight: tango
    toc_float:
      collapsed: false
      smooth_scroll: false
    css: style.css
    fig_width: 8
    fig_height: 3
date: "2024-08-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(data.table)
library(ggplot2)
library(ggpubr)
library(plotly)
library(magrittr)
library(dygraphs)
library(manipulateWidget)
library(crosstalk)
```

# The data

The unit of observation is a curve. The data is a collection of \(N\) curves \(\{X_1, \ldots, X_N\}\) that are realizations of a process $X$. Each curve $X_n$ is observed on the grid $t_1, t_2, \ldots, t_M$, where $M \geq 2$, with some measurement error $\varepsilon_{n,k}$:

$$
Y_{n,k} = X_n(t_k) + \varepsilon_{n,k},  \qquad n = 1, \ldots, N, \quad k = 1, \ldots, M.
$$
 
We assume that $\varepsilon_{n,k}$ is centered with variance $\mathbb{E}(\varepsilon_{n,k}^2) = \sigma^2(t_k)$.
 

**Example:** Daily Load Profile Curves of a Wind Farm

- **Curve per Day**: Each day is represented by a curve.
- **Observations per Day**: Every 10 minutes, an observation is recorded, resulting in 144 observations per curve.


# Prection using the BLUP

Let $n_0 \in \{1, \ldots, N\}$ and $t_0 \in \{t_1, \ldots, t_M\}$. The objective is to predict $X_{n_0}(t_0)$ using the curves $n_0$ and $n_0 - 1$. We use the BLUP, which stands for Best Linear Unbiased Predictor. Let us state its formula. Let


\begin{align*}
  \mathcal{Y}_{n_0,1} &=(Y_{n_0-1, 1},\ldots,Y_{n_0-1, M}, Y_{n_0, 1},\ldots,Y_{n_0, M})^\top \\
  \mathcal{M}_{n_0,1} &= (\mu(t_1),\ldots, \mu(t_M),\mu(t_1),\ldots, \mu(t_M))^\top \\
  \Sigma_{n_0} &= (\sigma(t_1),\ldots, \sigma(t_M),\sigma(t_1),\ldots, \sigma(t_M))^\top.
\end{align*}

The BLUP is
$$
  \mathfrak{X}_{n_0}(t_0) = \mu(t_0) + B_{n_0, 1}^\top (\mathcal{Y}_{n_0, 1}- \mathcal{M}_{n_0, 1}),
$$
where
$$
  B_{n_0, 1} = \operatorname{Var}_{M,T}(\mathcal{Y}_{n_0, 1})^{-1} \operatorname{Cov}_{M,T}\left(\mathcal{Y}_{n_0, 1},{X}_{n_0}(t_0)\right),
$$
\begin{equation}\label{def:var}
  \operatorname{Var}_{M,T} \left(\mathcal{Y}_{n_0,1}\right) = 
  \begin{pmatrix}
    G_0^{(n_0-1,n_0-1)} + \Sigma_{n_0-1} &  G_1^{(n_0-1,n_0)}\\
    G_1^{(n_0,n_0-1)}& G_0^{(n_0,n_0)}+ \Sigma_{n_0}
  \end{pmatrix}\in\mathbb R^{2 M}\times \mathbb R^{2 M},
\end{equation}

\begin{equation}\label{def:covar}
  \operatorname{Cov}_{M,T} \left(\mathcal Y_{n_0,1},  {X}_{n_0}(t_0)   \right) =
  \begin{pmatrix}
    \Gamma_1(t_1,t_0)\\
    \vdots\\
    \Gamma_1(t_M,t_0)\\
    \Gamma_0(t_1,t_0)\\
    \vdots\\
    \Gamma_0(t_M,t_0)
  \end{pmatrix}\in\mathbb R^{2 M},
\end{equation}
and 
$$
  G_\ell^{(n, n^\prime)} = \left(\Gamma_\ell(t_i,t_j)\right)_{1\leq i, j\leq M}.
$$

The mean and lag-$\ell$ (auto)covariance  estimator are

$$
  \widehat{\mu}_N(t_k) = \frac{1}{N} \sum_{n=1}^{N} Y_{n,k}, \qquad k=1, \ldots, M.
$$

$$
  \widehat{\Gamma}_\ell(t_i,t_k) = \frac{1}{N - \ell} \sum_{n=1}^{N - \ell} \{Y_{n,i} - \widehat{\mu}_N(t_i)\}\{Y_{n+\ell,j} - \widehat{\mu}_N(t_j)\}, \qquad i,j \in \{1, \ldots, M\} \quad \text{and} \quad \ell = 0, 1, \ldots, N-1.
$$


## Wind farm data application

We consider a single park, specifically the one with the fewest missing values in its production record. We assume the following:

- **Daily Curve Data**: Each day provides one curve with 144 records, recorded every ten minutes.
- **Functional Time Series (FTS)**: The collection of daily curves is considered as a functional time series.
- **Winter Curves**: Only curves from winter are used for checking stationarity.

To construct the Best Linear Unbiased Prediction (BLUP), we need the following:

- **Mean of the Process**: The average behavior of the process.
- **Covariance of the Process**: The variability and relationship between different points in the process.
- **Lag-$\ell$ Autocovariance of the Process**: The autocovariance for lags $\ell = 1, \ldots, N$.

```{r, echo=TRUE}
source("../R/model_blup.R")
dt_fts <- readRDS("../PrevProd_Eol/dt_fts.RDS")
```

```{r}
dt_fts
```


**predict $7/8$ of the last curve**
```{r, fig.width= 10, fig.height= 4, fig.align='center', echo=TRUE}
dt_res_blup <- get_blup_prev(
  data = dt_fts, 
  id_curve_to_pred = NULL,
  threshold = 1/8, 
  L = 1, 
  graph = TRUE)
```

**predict $3/4$ of the last curve**
```{r, fig.width= 10, fig.height= 4, fig.align='center', echo=TRUE}
dt_res_blup <- get_blup_prev(
  data = dt_fts, 
  id_curve_to_pred = NULL,
  threshold = 2/8, 
  L = 1, 
  graph = TRUE)
```

**predict $1/2$ of the last curve**
```{r, fig.width= 10, fig.height= 4, fig.align='center', echo=TRUE}
dt_res_blup <- get_blup_prev(
  data = dt_fts, 
  id_curve_to_pred = NULL,
  threshold = 4/8, 
  L = 1, 
  graph = TRUE)
```


**predict $1/4$ of the last curve**
```{r, fig.width= 10, fig.height= 4, fig.align='center', echo=TRUE}
dt_res_blup <- get_blup_prev(
  data = dt_fts, 
  id_curve_to_pred = NULL,
  threshold = 6/8, 
  L = 1, 
  graph = TRUE)
```

## PV data application

TODO

# Prection using FAR(1)

TODO : Introduce the method

## Wind farm data application

```{r, echo=TRUE}
# Predict the next curve
id_curve_to_prev <- dt_fts[, max(id_curve)]
data_for_comparison <- copy(dt_fts[id_curve == id_curve_to_prev])

# Predict
dt_far_prev <- get_far_prev(
  data = dt_fts[! id_curve == id_curve_to_prev],
  horizon = 1
)
```

```{r, fig.width= 7, fig.height= 4, fig.align='center'}
# plot prev
dt_graph <- rbind(
  data_for_comparison[, .("t" = tobs, "value" = X, Variable = "FC")],
  dt_far_prev[horizon == 1, .(t, "value" = prev, Variable = "Forecast")]
)
ggplot(dt_graph, aes(x = t, y = value, color = Variable, group = Variable)) +
  geom_line() +
  xlab("t") +
  ggtitle("Forecast vs True") + 
  ylab("") +
  scale_colour_manual(values = c("FC" = "#1B4F72", "Forecast" = "#A93226")) +
  theme_minimal() +
  theme(axis.title = element_text(size = 12),
        axis.title.x = element_text(size = 12, margin = margin(t = 10, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(size = 12, margin = margin(t = 10, r = 10, b = 0, l = 0)),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        legend.position = "bottom")
```



